\input{core_introduction}

\chapter{core. The Core Functionality}
\renewcommand{\curModule}{core}
\input{core_basic_structures}
\input{core_array_operations}
\input{core_dynamic_structures}
\input{core_drawing_functions}
\input{core_persistence}
\input{core_clustering_search}
\input{core_utilities_system_functions}


\chapter{imgproc. Image Processing}
\renewcommand{\curModule}{imgproc}
\input{imgproc_histograms}
\input{imgproc_image_filtering}
\input{imgproc_image_warping}
\input{imgproc_image_transform}
\input{imgproc_struct_shape_analysis}
\input{imgproc_planar_subdivisions}
\input{imgproc_motion_tracking}
\input{imgproc_feature_detection}
\input{imgproc_object_detection}

\chapter{features2d. Feature Detection and Descriptor Extraction}
\renewcommand{\curModule}{features2d}
\input{features2d_detection_description}
\input{features2d_common_detection_description}
\input{features2d_object_categorization}

\chapter{flann. Clustering and Search in Multi-Dimensional Spaces}
\renewcommand{\curModule}{flann}
\input{flann}

\chapter{objdetect. Object Detection}
\renewcommand{\curModule}{objdetect}
\input{objdetect}

\chapter{video. Video Analysis}
\renewcommand{\curModule}{video}
\input{video_motion_tracking}

\chapter{highgui. High-level GUI and Media I/O}
\renewcommand{\curModule}{highgui}
\input{highgui}
\ifPy     %Qt is for C and Cpp, so do nothing
\else
\input{highgui_qt}
\fi

\chapter{calib3d. Camera Calibration, Pose Estimation and Stereo}
\renewcommand{\curModule}{calib3d}
\input{calib3d}


\chapter{ml. Machine Learning}
\renewcommand{\curModule}{ml}
\input{ml}


